---
title: "[IPO1] Prepare data (and impute IPO)"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
---
  

```{r setup, cache=T, eval=T, echo=F}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
options(width=100)
```

<br>
<br>

## Read data

```{r, echo=T, eval=T, comment=""}
df_patient <- read.csv("./data/IPO_Beteg_TCI_SCL2_datjav.csv", header=T, sep=",") 
dim(df_patient)
table(df_patient$KOR2 < 18)  #KOR2=age
```
Unfortunately there is a 17-year old in the sample that has to be removed. 

```{r, echo=TRUE, eval=T, comment=""}
df_patient <- subset(df_patient, KOR2 > 17 | is.na(KOR2)) 
dim(df_patient) 
```

So this leaves 187 patients' data on 1399 variables for analysis.

<br>
<br>

# Missing data

Let's have a look at missing data structure of IPO to (hopefully) save some data by imputing. 
```{r, echo=TRUE, eval=T, comment="", fig.width=8, message=F, warning=F, fig.width=14}
library(VIM)
IPO_items <- df_patient[grep("^SZSZK", names(df_patient), value=T)]
aggr_plot <- aggr(IPO_items, col=c('navyblue','red'), numbers=T, sortVars=F, 
                  labels=names(IPO_items), cex.axis=0.9, cex.lab = 1.2, gap=1.2,
                  combined=F, 
                  ylab=c("Histogram of missing data","Pattern"))

```

87.7% of data are not missing any value (see plot on the right hand side), and 3.2% (so 6 participants) miss data on all IPO items. These will have to be excluded from the final database.  Other than this the distribution of missing data seems random to me. 
In terms of missing data of the items, Items 33 and 24 miss most data, although only 4.8% which isn't very much.  So we can proceed with data imputation. 

<br>

So we have 181 rows with SOME data and 6 rows with no data on IPO. 
<br>
Note the completely missing data rows at the end (last 6 rows). These will need to be omitted for imputation purposes.
```{r, echo=TRUE, eval=T, comment=""}
rowSums(is.na(IPO_items))
```


<br>
<br>

# Data imputation 

```{r, echo=TRUE, eval=T, comment="", message=F}
# So we want to get rid of rows which contain no data.
IPO_items_noBlankRow <- df_patient[grep("^SZSZK", names(df_patient), value=T)][!(rowSums(is.na(IPO_items)) == "83"),] 

# The rest can be inputed:
dim(IPO_items_noBlankRow) 
```

``` {r, echo=T, eval=T, results="hide", include=T, message=F}
# let's do the imputation 
library(mice)
IPO_items_imputed <- mice(IPO_items_noBlankRow, m=2, maxit=50, method="pmm", seed=1239)
```

```{r, echo=T, eval=T, message=F}
# and pick the 2nd imputed matrix as the final dataset which hopefully contains no missing data
IPO_items_complete <- complete(IPO_items_imputed, 2) 
dim(IPO_items_complete)
```

So, are there any missing data?
```{r, echo=TRUE, eval=T, comment=""}
rowSums(is.na(IPO_items_complete))
```
No more missing data here.

<br>

Finally, save the final, imputed version of IPO items.  
```{r, echo=T, eval=T, comment=""}
dim(IPO_items_complete)
write.csv(IPO_items_complete, file="./data/IPO_items_complete.csv")
```


