---
title: '[IPO2b] check IPO scales'
author: ''
date: ''
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
<br>

First of all, I'd like to check whether scales are normally distributed. This is important for the latent class analysis and for later, to assess group differences. 

```{r, eval=T, echo=TRUE, comment="", message=F, warning=F, fig.width=14, fig.height=8}

## Create the dataframe. 
IPO_items_complete <- read.csv(file="./data/IPO_items_complete.csv")
IPO_items_complete$X <- NULL # row index is not needed here
## we need the values numeric (instead of integer)
IPO_item_names <- colnames(IPO_items_complete)
  for(i in IPO_item_names) {
  IPO_items_complete[[i]] <- as.numeric(IPO_items_complete[[i]])
  }

## let's calculate the IPO scales
library(psych)
ipo.keys <- make.keys(IPO_items_complete, list(
  PPD = c(1:16), # primitive psychological defenses
  RT = c(17:37), # reality testing
  ID = c(38:57), # identity diffusion
  AGG = c(58:75), # aggression
  MV = c(6, 14, 17, 76:83) # moral values
  ))
IPO_scales_sum <- data.frame(scoreItems(ipo.keys, IPO_items_complete, totals=F)$scores)

## plot scale means
par(mfrow=c(3, 2), mar=c(4, 4, 2, 1))
hist(IPO_scales_sum$PPD, main="Primitive Psychological Defenses", xlab = "", breaks = 40)
hist(IPO_scales_sum$RT, main="Reality Testing", xlab = "", breaks = 40)
hist(IPO_scales_sum$ID, main="Identity Diffusion", xlab = "", breaks = 40)
hist(IPO_scales_sum$AGG, main="Aggression", xlab = "", breaks = 40)
hist(IPO_scales_sum$MV, main="Moral Values", xlab = "", breaks = 40)
```

The Aggression scale appears to be left-scewed, but the other scales can be treated as approximately normally distributed. 

Let's check normal distribution **statistically** using the Shapiro-Wilk test (W) and Anderson-Darling test (AD). 

```{r, eval=T, echo=T}
library(nortest)

df_IPOnorm <- data.frame("W" = numeric(), "p_W" = numeric(), 
                         "AD" = numeric(), "p_AD" = numeric())

for (i in 1:ncol(IPO_scales_sum)) {
  columns <- names(IPO_scales_sum[i])
  shapiroResult <- shapiro.test(IPO_scales_sum[,i])
  ADResult <- ad.test(IPO_scales_sum[,i])
  newRow <- list(shapiroResult[1], 
                 shapiroResult[2], 
                 ADResult[1], 
                 ADResult[2])
  df_IPOnorm[names(IPO_scales_sum[i]), ] <- newRow
  }

print(df_IPOnorm)

``` 
PPD and RT appear to be normally distributed, but not the other scales. This will not affect the Latent Class Analysis, but violates the assumption of ANOVA. 


