---
title: '[IPO2a] check IPO items'
author: ''
date: ''
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

#### In .pdf format to reduce file size & keep readability. 

<br>
<br>

First of all, let's have a look at the stucture of IPO items.
```{r, echo=TRUE, eval=T, comment="", fig.width=18, fig.height=15}
IPO_items_complete <- read.csv(file="./data/IPO_items_complete.csv")
IPO_items_complete$X <- NULL

library(corrplot)
IPO_items_complete_M <- cor(IPO_items_complete)
corrplot(IPO_items_complete_M, method="circle") 
```

As evident from this figure, there are no reversely coded items. This is good news, because although the instrument was not designed to be used as a severity scale (by using the total score), each item measures the same construct: personality integrity. In addition, judging from the pattern it seems that items seem to load on multiple factors. But we'll check this later. 
<br>
<br>

## Checking the IPO items

```{r, echo=T, eval=T, comment="", message=F, warning=F, fig.width=18, fig.height=12}
library(reshape2)
IPO_items_complete_melt <- melt(IPO_items_complete)

library(ggplot2)
ggplot(data=IPO_items_complete_melt, aes(x=value)) +
  facet_wrap(~variable, scales="free_x") +
  geom_histogram(binwidth=1)
```

Many items are not normally distributed (they are left skewed), so items will be treated as categorical variables in CFA. This may be a sign that IPO is more sensitive in the middle and upper severity of personality organisation than in the "low" severity (or well-organised) group. Alternatively, it may indicate that this is a low to medium severity sample in terms of personality organisation. 

<br>

In terms of data analysis, however, moderately skewed items will not be a problem, because the unit of data analysis on the IPO is the scales, not the items. And scales are approximately normally distributed (as you'll see in the next file). 

