---
title: '[IPO6b] Corrplot of scales (SCL-90, STAXI and DES)'
author: ''
date: ''
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, eval=T}
knitr::opts_chunk$set(echo = TRUE, fig.width = 15, fig.height = 15)
options(width=90)
```

<br>
<br>

Data are still the same: 
```{r, cache=T, echo=F, message=F, warning=F, error=F}
df <- read.csv(file="/home/aniko/R/IPO/data/df_final.csv")
dim(df) # 181 rows and 585 (created an extra ID (=X) variable)

library(plyr)
df <- plyr::rename(df, c("X"="id"))

df_names <- colnames(df)
for(i in df_names) {
  df[[i]] <- as.numeric(df[[i]])
}

df$gender <- as.factor(df$gender)
df$education <- as.factor(df$education)
df$lca_class <- as.factor(df$lca_class)

# Now get the databases. 
df_staxi <- df[grep("^STAX", names(df), value=T)]
df_des <- df[grep("^DES", names(df), value=T)]
df_scl <- df[grep("^SCL", names(df), value=T)]

# Prepare the scales. 
library(psych)
scl.keys <- make.keys(df_scl, list(
  som = c(1, 12, 27, 4, 40, 42, 48, 49, 52, 53, 56, 58), # Somatization
  ocd = c(10, 28, 3, 38, 45, 46, 51, 55, 65, 9), # Obsession
  hos = c(11, 24, 63, 67, 74), # Hostility
  pho = c(13, 25, 47, 50, 70, 75, 82), # Phobic Anxiety
  dep = c(14, 15, 20, 22, 26, 29, 30, 31, 32, 5, 54, 71, 79), # Depression
  psy = c(16, 35, 62, 7, 77, 84, 85, 87, 88, 90), # Psychoticism
  anx = c(17, 2, 23, 33, 39, 57, 72, 78, 80, 86), # Anxiety
  par = c(18, 43, 68, 76, 8, 83), # Paranoia
  int = c(21, 34, 36, 37, 41, 6, 61, 69, 73), # Interpersonal Sensitivity
  SUM = c(1:90)
))

staxi.keys <- make.keys(df_staxi, list(  # note that I translated the scale names from Hun. Original names need to be checked
  state = c(1:25), # State Anger
  expr = c(26:57), # Anger expression and anger control
  ang_fe = c(1, 2, 3, 6, 10), # Feeling of anger
  ang_ve = c(4, 9, 12, 13, 15), # Verbal expression of anger
  ang_ph = c(5, 7, 8, 11, 14), # Bodily (physical) expression of anger
  ang_te = c(16, 17, 18, 21), # Angry temperament
  ang_re = c(19, 20, 23, 25), # Anger reaction
  ang_expr = c(27, 31, 35, 39, 43, 47, 51, 55), # Expression of anger 
  ang_rep = c(29, 33, 37, 41, 45, 49, 53, 57), # Repression of anger
  ang_expr_ctrl = c(26, 30, 34, 38, 42, 46, 50, 54), # Controlling the expression of anger
  ang_rep_ctrl = c(28, 32, 36, 40, 44, 48, 52, 56) # Controlling the repression of anger
  ))

des.keys <- make.keys(df_des, list(
  amn = c(3, 4, 5, 6, 8, 10, 25, 26), # Amnesia
  abs = c(2, 14, 15, 16, 17, 18, 20, 22, 23), # Absorption in imagination
  depers = c(7, 11, 12, 13, 27, 28) # Depersonalisation-derealisation
))



# Let's calculate the scales. 
SCL_scores <- scoreItems(scl.keys, df_scl, totals=F, missing=TRUE, impute="median")
# print(SCL_scores$alpha, digits=2)
STAXI_scores <- scoreItems(staxi.keys, df_staxi, totals=F, missing=TRUE, impute="median") 
# print(STAX_scores$alpha, digits=2)
# IPO_scales_sum  # don't forget we also had this. 
DES_scores <- scoreItems(des.keys, df_des, totals=F, missing=TRUE, impute="median")
#print(DES_scores$alpha, digits=2)
```

<br>

# SCL-90

First of all, there does not seem to be any reversed item. I am also glad to see that this is a more-or-less homogeneous instrument given the positive (r>0.1) correlations between items that are designed to measure the same concept (symtom severity). 

```{r, eval=T, echo=F}
library(corrplot)
scl_items <- df[grep("^SCL[1-90]+", names(df), value=T)]
M_scl <- cor(scl_items, use="complete.obs")
corrplot(M_scl, method="circle") #no reversed items!
```

\pagebreak

```{r, eval=T, echo=T}
library(psych)
SCL_scores$alpha
alpha(scl_items)[1]
```

Cronbach alpha for the total is acceptable (0.97), scale alphas range between 0.76 and 0.91. Items correlate with each other at a level of r=0.31 on average, and the overall mean (1.85) is centered around the mean of response alternatives (0-4).

<br>

#STAXI
STAXI is more heterogenous than SCL-90. There are items which correlate negatively with one another, but according to the initial publication, there are no reversely coded items. You can also clearly see the two "squares" in the matrix: this is an indicaton that there are at least two distinct factors among the items (and indeed this is the case, although there are many more factors, not just 2).


```{r, eval=T, echo=F}
staxi_items <- df[grep("^STAX+", names(df), value=T)]
M_staxi <- cor(staxi_items, use="complete.obs")
corrplot(M_staxi, method="circle") #there seem to be a few reversed items here. 
```

```{r, eval=T, echo=T, warning=F, error=T}
library(psych)
STAXI_scores$alpha
alpha(staxi_items)[1]
```

Cronbach alphas are all the the acceptable level (all >= 0.75), but the mean correlation among items is "only" 0.12, because of the many negative correlational coefficients. The scale is by no means suitable as a "total" anger measure, but it wasn't designed to be used as total either. When we have a closer look at the negative correlations, it becomes evident that "the expression of anger" items are the ones which correlate negatively with "anger repression" items, which makes sense. 

<br>

#DES

DES items are homogeneous as well, which "makes sense", given that the instrument was developped to measure dissociation (on 3 separate but related dimensions).

``````{r, eval=T, echo=F}
des_items <- df[grep("^DES", names(df), value=T)]
M_des <- cor(des_items, use="complete.obs")
corrplot(M_des, method="circle") 
```

No reversed items based on visual inspection.

\pagebreak


```{r, eval=T, echo=T, warning=F, error=T}
library(psych)
DES_scores$alpha
alpha(des_items)[1]
```

Cronbach alphas are again excellent (>0.78), with the mean correlation around 0.36. 