---
title: "[IPO7] SCL DES STAXI  plot with LCA"
author: ""
date: ""
output: 
  html_document: 
    code_folding: hide
---


```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo=T, comment = "")
```

<br>
<br>

####*Note that you can unhide the code by clicking on the button **"Code"** on the right.* 
<br>
<br>

Data is still the same: 

```{r, cache=T, echo=F, message=F, warning=F, error=F}
df <- read.csv(file="/home/aniko/R/IPO/data/df_final.csv")
dim(df) # 181 rows and 585 (created an extra ID (=X) variable)

library(plyr)
df <- plyr::rename(df, c("X"="id"))

df_names <- colnames(df)
for(i in df_names) {
  df[[i]] <- as.numeric(df[[i]])
}

df$gender <- as.factor(df$gender)
df$education <- as.factor(df$education)
df$lca_class <- as.factor(df$lca_class)

# Now get the databases. 
df_staxi <- df[grep("^STAX", names(df), value=T)]
df_des <- df[grep("^DES", names(df), value=T)]
df_scl <- df[grep("^SCL", names(df), value=T)]

# Prepare the scales. 
library(psych)
scl.keys <- make.keys(df_scl, list(
  som = c(1, 12, 27, 4, 40, 42, 48, 49, 52, 53, 56, 58), # Somatization
  ocd = c(10, 28, 3, 38, 45, 46, 51, 55, 65, 9), # Obsession
  hos = c(11, 24, 63, 67, 74), # Hostility
  pho = c(13, 25, 47, 50, 70, 75, 82), # Phobic Anxiety
  dep = c(14, 15, 20, 22, 26, 29, 30, 31, 32, 5, 54, 71, 79), # Depression
  psy = c(16, 35, 62, 7, 77, 84, 85, 87, 88, 90), # Psychoticism
  anx = c(17, 2, 23, 33, 39, 57, 72, 78, 80, 86), # Anxiety
  par = c(18, 43, 68, 76, 8, 83), # Paranoia
  int = c(21, 34, 36, 37, 41, 6, 61, 69, 73), # Interpersonal Sensitivity
  SUM = c(1:90)
))

staxi.keys <- make.keys(df_staxi, list(  # note that I translated the scale names from Hun. Original names need to be checked
  state = c(1:25), # State Anger
  expr = c(26:57), # Anger expression index
  ang_fe = c(1, 2, 3, 6, 10), # Feeling of anger
  ang_ve = c(4, 9, 12, 13, 15), # Verbal expression of anger
  ang_ph = c(5, 7, 8, 11, 14), # Bodily (physical) expression of anger
  ang_te = c(16, 17, 18, 21), # Angry temperament
  ang_re = c(19, 20, 23, 25), # Anger reaction
  ang_expr = c(27, 31, 35, 39, 43, 47, 51, 55), # Expression of anger 
  ang_rep = c(29, 33, 37, 41, 45, 49, 53, 57), # Repression of anger
  ang_expr_ctrl = c(26, 30, 34, 38, 42, 46, 50, 54), # Controlling the expression of anger
  ang_rep_ctrl = c(28, 32, 36, 40, 44, 48, 52, 56) # Controlling the repression of anger
  ))

des.keys <- make.keys(df_des, list(
  amn = c(3, 4, 5, 6, 8, 10, 25, 26), # Amnesia
  abs = c(2, 14, 15, 16, 17, 18, 20, 22, 23), # Absorption in imagination
  depers = c(7, 11, 12, 13, 27, 28) # Depersonalisation-derealisation
))



# Let's calculate the scales. 
SCL_scores <- scoreItems(scl.keys, df_scl, totals=F, missing=TRUE, impute="median")
# print(SCL_scores$alpha, digits=2)
STAXI_scores <- scoreItems(staxi.keys, df_staxi, totals=F, missing=TRUE, impute="median") 
# print(STAX_scores$alpha, digits=2)
# IPO_scales_sum  # don't forget we also had this. 
DES_scores <- scoreItems(des.keys, df_des, totals=F, missing=TRUE, impute="median")
#print(DES_scores$alpha, digits=2)
```

<br>

# Symptom severity (SCL-90):
``` {r, message=F, warning=F, eval=T, fig.width=12, fig.height=7}
scl_scores <- data.frame(SCL_scores$scores)
scl_scores$SUM <- NULL
scl_scores$lca_class <- df$lca_class

library(reshape2)
scl_scores_lca_melt <- melt(scl_scores, id="lca_class") 

library(Rmisc)
scl_plot <- summarySE(scl_scores_lca_melt, measurevar=c("value"), groupvars=c("lca_class", "variable"), na.rm=T)

scl_plot$lca_class <- factor(scl_plot$lca_class, 
                             levels=c("2", "3", "1"), 
                             labels=c("Disintegrated", "Moderately integrated", "Well-integrated"))

library(ggplot2)
pd <- position_dodge(width=0.3)
addline_format <- function(x, ...){
  gsub('\\s', '\n', x)
  }

ggplot(scl_plot, aes(x=variable, y=value, 
                              colour=lca_class, group=lca_class)) + 
  coord_cartesian(ylim=c(0,4)) +
  geom_errorbar(aes(ymin=value-ci, ymax=value+ci), 
                color="black", width=.3, size=1, 
                position = pd) +
  geom_line(size=1.5, position=pd) +
  geom_point(size=1.5, position=pd) +
  scale_colour_manual(values = c("red","blue", "green4")) +
  theme_minimal(base_size=20) +
  labs(y="Scale mean", x="", color="", title = "") + 
  scale_x_discrete(labels=addline_format(c("som" = "Somatization", 
                            "ocd" = "Obsession", 
                            "hos" = "Hostility", 
                            "pho" = "Phobic Anxiety", 
                            "dep" = "Depression", 
                            "psy" = "Psychoticism",
                            "anx" = "Anxiety", 
                            "par" = "Paranoia", 
                            "int" = "Interpersonal Sensitivity"), 
                            breaks=unique(g_scl$variable))) +
  theme(legend.position= "top", 
        legend.key.size = unit(3, "cm"),
        axis.text.x=element_text(colour="black"))
```

Group differences on SCL-90 with 95% confidence interval. Clearly, the Disintegrated group scores higher on distress. Note that the extent of personality integration (as measured with IPO, depicted earlier) corresponds to the extent of psychiatric symptoms between latent classes. 

<br>
<br>

# State-Trait Expression Inventory (STAXI): 
``` {r, message=F, warning=F, fig.width=14, fig.height=8}
 staxi_scores <- data.frame(STAXI_scores$scores)
 staxi_scores$lca_class <- df$lca_class
 # head(stax_scores_lca)
 staxi_scores_lca_melt <- melt(staxi_scores, id="lca_class") 
 
library(Rmisc)
staxi_plot <- summarySE(staxi_scores_lca_melt, measurevar=c("value"),
                         groupvars=c("lca_class", "variable"), na.rm=T)
staxi_plot$lca_class <- factor(staxi_plot$lca_class, levels=c("2", "3", "1"), 
                              labels=c("Disintegrated", "Moderately integrated", "Well-integrated"))

library(ggplot2)
pd <- position_dodge(width=0.3)
addline_format <- function(x, ...){
  gsub('\\s', '\n', x)
  }

ggplot(staxi_plot, aes(x=variable, y=value, 
                      colour=lca_class, group=lca_class)) + 
  coord_cartesian(ylim=c(0.5,4)) +
  geom_errorbar(aes(ymin=value-ci, ymax=value+ci), 
                color="black", width=.3, size=1, position = pd) +
  geom_line(size=1.5, position=pd) +
  geom_point(size=1.5, position=pd) +
  theme_minimal(base_size=20) +
  scale_colour_manual(values = c("red", "blue", "green4")) +
  labs(y="Scale mean", x="", color="", title = "")  +
  scale_x_discrete(labels=addline_format(c(
    "state" = "State Anger",  
    "expr" = "Anger expression index",
    "ang_fe" = "Feeling angry", 
    "ang_ve" = "Feel like expressing anger verbally",  
    "ang_ph" = "Feel like expressing anger physically",  
    "ang_te" = "Trait anger", 
    "ang_re" = "Angry temperament", 
    "ang_expr" = "Anger expression - out",  
    "ang_rep" = "Anger expression - in", 
    "ang_expr_ctrl" = "Anger control - out", 
    "ang_rep_ctrl" = "Anger control - in"), 
    breaks=unique(staxi_plot$variable))) + 
  theme(legend.position="top", 
        legend.key.size = unit(3, "cm"), 
        axis.text.x=element_text(colour="black"))

```

This is the maximum number of scales calculated from STAX items and yes, it is perhaps too many... But: as a trend, Disintegrated individuals seem to be more "angry" than the other 2 groups. The difference is the largest in scales of temperament, reaction and repression. This resonates with the literature that says that patients suffering from borderline (or other severe) personality disorder have difficulties controlling their anger despite having higher tendencies of anger reaction on the biological level (enlarged amygdala and weakened frontal lobe connections to allow for controlling). 

<br>
<br>

# Dissociative Experiences Scale (DES): 
``` {r, message=F, warning=F}
des_scores <- data.frame(DES_scores$scores)
des_scores$lca_class <- df$lca_class
des_scores_lca_melt <- melt(des_scores, id="lca_class") 

library(Rmisc)
des_plot <- summarySE(des_scores_lca_melt, measurevar=c("value"), groupvars=c("lca_class", "variable"), na.rm=T)
des_plot$lca_class <- factor(des_plot$lca_class, 
                             levels=c("2", "3", "1"), 
                            labels=c("Disintegrated", "Moderately integrated", "Well-integrated"))

library(ggplot2)
pd <- position_dodge(width=0.3)
addline_format <- function(x, ...){
  gsub('\\s', '\n', x)
}

ggplot(des_plot, aes(x=variable, y=value, colour=lca_class, group=lca_class)) + 
  coord_cartesian(ylim=c(0,60)) +
  geom_errorbar(aes(ymin=value-ci, ymax=value+ci), 
                color="black", width=.3, size=1, position = pd) +
  geom_line(size=1.5, position=pd) +
  geom_point(size=1.5, position=pd) +
  theme_minimal(base_size=20) +
  scale_colour_manual(values = c("red", "blue", "green4")) +
  labs(y="Mean %", x="", color="", 
       title = "")  +
  scale_x_discrete(labels=addline_format(c(
    "amn" = "Amnesia",
    "abs" = "Absorption in imagination",
    "depers" = "Depersonalisation- Derealisation"
   ), 
    breaks=unique(des_plot$variable))) +
  theme(legend.position="top", 
        legend.key.size = unit(1, "cm"), 
        axis.text.x=element_text(colour="black"))
```

Although psychotic patients were excluded from the sample, clearly, severly personality disordered individuals display more dissociative experiences than others. This is especially evident in the "Absorption in imagination" scale, which indicates vivid projective mechanisms as a way of experiencing reality. 


