---
title: "[IPO8] diagnoses (Axis1 & Axis2)"
author: ""
date: ""
output: 
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, comment = "")
```

<br>

####*Note that you can see the code by clicking on the **"Code"** button on the right. *

```{r, cache=T, echo=F, message=F, warning=F, error=F}
df <- read.csv(file="/home/aniko/R/IPO/data/df_final.csv")
# dim(df) # 181 rows and 606 (created an extra ID (=X) variable)

library(plyr)
df <- plyr::rename(df, c("X"="id"))

df_names <- colnames(df)
for(i in df_names) {
  df[[i]] <- as.numeric(df[[i]])
}

df$gender <- as.factor(df$gender)
df$education <- as.factor(df$education)
df$lca_class <- as.factor(df$lca_class)

# Now get the databases. 
df_axis1_dg <- df[grep("^S1DG[1-9]$|S1DG10|S1DG11|
                      |S1DG12|S1DG13|S1DG16|S1DG17|S1DG18|S1DG19|S1DG20|S1DG21|S1DG22|
                       |S1DG25|S1DG26|S1DG27|S1DG28|S1DG29|S1DG30|S1DG31|S1DG32|S1DG33|
                       |S1DG40|S1DG41|S1DG42|S1DG43|S1DG44|S1DG45|S1DG46|S1DG47|S1DG49|
                       |S1DG53|S1DG54", names(df), value=T)]
df_axis2_dg <- data.frame(df[grep("S7_32|S15_41|S24_50|S32_58|S40_66|S48_74|
                       |S59_84SZ|S65_92|S72_101|S89_111|S104_121|S119_139", names(df), value=T)])
df_axis2_sever <- data.frame(df[grep(
                              "S1_25|S2_26|S3_27|S4_28|S5_29|S6_30|S7_31|S8_33|S9_34|
                              |S10_35|S11_36|S12_37|S13_38|S14_39|S16_42|S17_43|
                              |S18_44|S19_45|S20_46|S21_47|S22_48|S23_49|^S24|S25_51|
                              |^S26|S27_52|S28_53|S29_54|S30_55|S31_56|S32_57|S33_59|
                              |S34_60|S35_61|S36_62|S37_63|S38|S39_64|S40_65|S41_67|
                              |S42_68|S43_69|S44_70|S45_71|S46|S47_72|S48_72|S49_75|
                              |S50|S51|S52_76|S53|S54|S55_77|S56|S57|S57_78|S57_79|S57_80|
                              |S57_81|S58_82|S59_84$|S60_85|S61_86|S62_87|S63_88|
                              |S63_89|S64_90|S65_91|S66_93|S67_94|S68_95|S69_96|S69_97|
                              |S70_98|S71_99|S72_100|S73_102|S74|S75_103|S76|S77_104|
                              |S78|S79_105|S80_106|S81_107|S82|S83|S84|S85_108|S86|S87_109|
                              |S88|S89_110|S90_112|S91_113|S92_114|S93|S94|S95|
                              |S96_115|S97_116|S98|S99_117|S100_118|S101_119|S102|S103|
                              |S104_120|S105_122|S105_123|S106_124|S107_125|S108_126|
                              |S110_128|S111_129|S112_130|S113_131|S114_132|S115_133|
                              |S116_134|S117_135|S118_136|S119_137|S119_138|S119_139", 
                            names(df), value=T
                              )])

# Prepare the scales. 
library(psych)

axis1_dg.keys <- make.keys(df_axis1_dg, list(
  axis1_dg_sum = c(1:40)
))

# df_axis1_dg$S1DG29  # there were no transsexual participants in this sample so this item was removed
# df_axis1_dg$S1DG25
# df_axis1_dg$S1DG20
# df_axis1_dg$S1DG21
# df_axis1_dg$S1DG30
# df_axis1_dg$S1DG33
## all of the above items were excluded because there were no such diagnoses (all of them were items to detect psychotic symptoms)


axis1_catego.keys <- make.keys(df_axis1_dg, list(
  anx = c("S1DG8", "S1DG9", "S1DG10", "S1DG11", "S1DG12", "S1DG16", 
          "S1DG22", "S1DG41", "S1DG54"),
  obs = c("S1DG5"), 
  depr = c("S1DG4", "S1DG43"), 
  bipol = c("S1DG31", "S1DG27", "S1DG28"), 
  # psych = c("S1DG25", "S1DG20", "S1DG21", "S1DG30", "S1DG33"), 
  somatof = c("S1DG17", "S1DG45", "S1DG46"), 
  dissoc = c("S1DG18"), 
  sud = c("S1DG7", "S1DG32", "S1DG40", "S1DG42", "S1DG44", "S1DG53"), 
  ED_imp = c("S1DG2", "S1DG3", "S1DG13"), 
  ED_rest = c("S1DG1") 
  # gend_id = c("S1DG29")
          ))
  

## let's recode the diagnoses so that only 3=positive, 2 and 1 == negative diagnosis
axis2_dg_names <- names(df_axis2_dg)
for(i in axis2_dg_names) {
  df_axis2_dg[[i]][df_axis2_dg[[i]] == "1"] <- "0"
  df_axis2_dg[[i]][df_axis2_dg[[i]] == "2"] <- "0"
  df_axis2_dg[[i]][df_axis2_dg[[i]] == "3"] <- "1"
}

for(i in axis2_dg_names) {
df_axis2_dg[[i]] <- as.numeric(df_axis2_dg[[i]])
}

axis2_dg.keys <- make.keys(df_axis2_dg, list(
  axis2_dg_sum = c(1:12)
))

## we also need the information how many personality disorder symptom was actually positive. 
## so we'll need to record these in the following way: 3=positive (1), 2 and 1 = negative (0) symptom
axis2_sever_names <- colnames(df_axis2_sever)
for(i in axis2_sever_names) {
  df_axis2_sever[[i]][df_axis2_sever[[i]] == "1"] <- "0"
  df_axis2_sever[[i]][df_axis2_sever[[i]] == "2"] <- "0"
  df_axis2_sever[[i]][df_axis2_sever[[i]] == "3"] <- "1"
}

for(i in axis2_sever_names) {
  df_axis2_sever[[i]] <- as.numeric(df_axis2_sever[[i]])
}

axis2_sever.keys <- make.keys(df_axis2_sever, list(
  axis2_sever_sum = c(1:127)
))

# Let's calculate the scales. 

axis1_dg_scores <- scoreItems(axis1_dg.keys, df_axis1_dg, totals=F, missing=T, impute="median")
axis1_dg_scores_total <- scoreItems(axis1_dg.keys, df_axis1_dg, totals=T, missing=T, impute="median")
# note that items 19, 20, 21, 25, 26, 29, 30 and 33 had no variance so they were deleted from keys.
axis1_catego <- scoreItems(axis1_catego.keys, df_axis1_dg, totals=F, missing=T, impute="median")
axis1_catego_total <- scoreItems(axis1_catego.keys, df_axis1_dg, totals=T, missing=T, impute="median")
# print(axis1_catego, digits=2)
axis2_dg_scores <- scoreItems(axis2_dg.keys, df_axis2_dg, totals=T)
# axis2_dg_scores$scores
axis2_sever_scores <- scoreItems(axis2_sever.keys, df_axis2_sever, totals=T, missing=T, impute="median")
# note that S111_129 had no variance (no positive cases in this sample) so was deleted from keys. 
```

<br>
<br>

<center> <h1>Axis 1</h1> </center>

<br>

### Number of diagnoses present per latent class:
``` {r, warning=F, message=F, eval=T}
axis1_dg_scores <- data.frame(axis1_dg_scores$scores)
axis1_dg_scores$lca_class <- df$lca_class
# head(axis1_dg_scores)
axis1_dg_scores$lca_class <- factor(axis1_dg_scores$lca_class, levels=c("3", "1", "2"), 
                              labels=c("Well-integrated", "Moderately integrated", "Disintegrated"))

library(ggplot2)
ggplot(axis1_dg_scores, aes(x=lca_class, y=axis1_dg_sum, fill=lca_class)) +
  geom_boxplot(color="black", lwd=1) +
  coord_cartesian(ylim=c(-0.02,0.2)) +
  theme_minimal(base_size=20) +
  scale_fill_manual(values=c("green4", "blue", "red")) +
  theme(legend.position="none", axis.text.x=element_text(colour="black")) +
  labs(y="Mean number of Axis I. diagnoses", x="", color="", 
       title = "") +
  theme(legend.position="none")
```

Notice that the disintegrated group has slightly more Axis I. diagnoses but this isn't a significant difference. 

<br>
<br>
<br>
<br>
<br>

### Diagnoses according to **categories**:
``` {r, eval=T, message=F, warning=F, fig.width=12, fig.height=8}
df_axis1_categories <- data.frame(axis1_catego$scores)
df_axis1_categories$lca_class <- df$lca_class 
# head(df_axis1_categories)

library(reshape2)
df_axis1_categories_melt <- melt(df_axis1_categories, id="lca_class") 

library(Rmisc)
axis1_categories_plot <- summarySE(df_axis1_categories_melt, 
                                   measurevar=c("value"), 
                                   groupvars=c("lca_class", "variable"), na.rm=T)

axis1_categories_plot$lca_class <- factor(axis1_categories_plot$lca_class, 
                                          levels=c("2", "3", "1"), 
                                          labels=c("Disintegrated", "Moderately integrated", "Well-integrated"))

library(ggplot2)
pd <- position_dodge(width=0.3)
addline_format <- function(x, ...){
  gsub('\\s', '\n', x)
  }

ggplot(axis1_categories_plot, 
                             aes(x=variable, y=value, 
                                 colour=lca_class, group=lca_class)) + 
  coord_cartesian(ylim=c(0,0.35)) +
  geom_errorbar(aes(ymin=value-ci, ymax=value+ci), 
                color="black", width=.3, size=1, position = pd) +
  geom_line(size=1.5, position=pd) +
  geom_point(size=1.5, position=pd) +
  theme_minimal(base_size=20) +
  theme(axis.text.x=element_text(angle=75, hjust=1)) +
  scale_colour_manual(values = c("red", "blue", "green4")) +
  labs(y="Scale mean", x="", color="", 
       title = "")  +
  scale_x_discrete(labels=addline_format(c(
  "anx" = "Anxiety Disorders",
  "obs" = "Obsessive Disorder", 
  "depr" = "Depressive Disorders", 
  "bipol" = "Bipolar Disorders", 
  "somatof" = "Somatoform Disorders", 
  "dissoc" = "Dissociative Disorder",  
  "sud" = "Substance- Related Disorders", 
  "ED_imp" = "Impulsive Eating Disorders", 
  "ED_rest" = "Restrictive Anorexia Nervosa"),
  breaks=unique(axis1_categories_plot$variable))) +
  theme(legend.position = "top", legend.key.size= unit(4, "cm"), axis.text.x=element_text(colour="black"))

```

There are no substantial differences between groups when using the categories which describe Axis I. symptoms. 
<br>
<br>
<br>
<br>
<br>

<center> <h1>Axis 2 (personality disorders)</h1> </center>

<br>
<br>


### Number of diagnoses:
```{r, eval=T}

axis2_dg_scores <- data.frame(axis2_dg_scores$scores)
axis2_dg_scores$lca_class <- df$lca_class

# head(axis2_dg_scores)
axis2_dg_scores$lca_class <- factor(axis2_dg_scores$lca_class, levels=c("1", "3", "2"), 
                                        labels=c("Well-integrated", "Moderately integrated", "Disintegrated"))

library(ggplot2)
ggplot(axis2_dg_scores, aes(x=lca_class, y=axis2_dg_sum, fill=lca_class)) +
  geom_boxplot(color="black", lwd=1) +
  coord_cartesian(ylim=c(-1,8)) +
  theme_minimal(base_size=20) +
  scale_fill_manual(values=c("green4", "blue", "red")) +
  theme(legend.position="none") +
  labs(y="Mean number of Axis II. diagnoses", x="", colour="", 
       title = "") +
  theme(legend.position = "none", axis.text.x=element_text(colour="black"))
```

Although the trend is clearer on Axis 2 (personality disorders), the difference between groups is probably not significant given the overlaping confidence intervals. 

<br>
<br>
<br>
<br>
<br>

### Personality disorders **one by one**:
```{r, eval=T, fig.width=12, fig.height=8}
df_axis2_onebyone <- mutate(df_axis2_dg)
df_axis2_onebyone$lca_class <- df$lca_class 

library(reshape2)
df_axis2_onebyone_melt <- melt(df_axis2_onebyone, id="lca_class") 

library(Rmisc)
axis2_onebyone_plot <- summarySE(df_axis2_onebyone_melt, 
                                 measurevar=c("value"), 
                                 groupvars=c("lca_class", "variable"), na.rm=T)

# head(axis2_onebyone_plot)

axis2_onebyone_plot$lca_class <- factor(axis2_onebyone_plot$lca_class, 
                                        levels=c("2", "3", "1"), 
                                        labels=c("Disintegrated", 
                                                 "Moderately integrated", 
                                                 "Well-integrated"))
library(ggplot2)
pd <- position_dodge(width=0.3)
addline_format <- function(x, ...){
  gsub('\\s', '\n', x)
  }

ggplot(axis2_onebyone_plot, aes(x=variable, y=value, colour=lca_class, group=lca_class)) + 
  coord_cartesian(ylim=c(0,1)) +
  geom_errorbar(aes(ymin=value-ci, ymax=value+ci), color="black", width=.3, size=1, position = pd) +
  geom_line(size=1.5, position=pd) +
  geom_point(size=1.5, position=pd) +
  theme_minimal(base_size=20) +
  theme(axis.text.x=element_text(angle=75, hjust=1)) +
  scale_colour_manual(values = c("red", "blue", "green4")) +
  labs(y="Probability (0-1)", x="", color="", 
       title = "")  +
  scale_x_discrete(labels=addline_format(c(
    "S7_32" = "Avoidant",
    "S15_41" = "Dependent",
    "S24_50" = "Obsessive- Compulsive",
    "S32_58" = "Passive- Agressive",
    "S40_66" = "Depressive",
    "S48_74" = "Paranoid",
    "S59_84SZ" = "Schizotypal",
    "S65_92" = "Schizoid",
    "S72_101" = "Histrionic",
    "S89_111" = "Narcissistic",
    "S104_121" = "Borderline",
    "S119_139" = "Antisocial"),
    breaks=unique(axis2_onebyone_plot$variable))) +
  theme(legend.position ="top", legend.key.size=unit(5, "cm"), axis.text.x=element_text(colour="black"))

```

However, when personality disorders are "expanded", it is clear that latent class differences are driven by borderline and depressive personality disorder. 

<br>
<br>
<br>
<br>
<br>

### Number of severe (coded as 3) symptoms:
``` {r, eval=T}

axis2_sever_scores <- data.frame(axis2_sever_scores$scores)
axis2_sever_scores$lca_class <- df$lca_class

axis2_sever_scores$lca_class <- factor(axis2_sever_scores$lca_class, levels=c("1", "3", "2"), 
                                        labels=c("Well-integrated", "Moderately integrated", "Disintegrated"))

library(ggplot2)
ggplot(axis2_sever_scores, aes(x=lca_class, y=axis2_sever_sum, fill=lca_class)) +
  geom_boxplot(color="black", lwd=1) +
  coord_cartesian(ylim=c(0, 55)) +
  theme_minimal(base_size=18) +
  scale_fill_manual(values=c("green4", "blue", "red")) +
  theme(legend.position="none") +
  labs(y="Number of Axis II. symptoms met (0-127)", x="", color="", 
       title = "") +
  theme(legend.position = "none", axis.text.x=element_text(colour="black"))

```

Again, it appears as a clear trend that the Disintegrated group has more "severe" personality disorder symptoms, this group difference in itself is not significant given the large variance. 
