---
title: "[IPO9] Calc1 (descriptive statistics)"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
options(width = 100)
```

<br>
<br>

Data are the same: 

```{r, cache=T, echo=F, message=F, warning=F, error=F}
df <- read.csv(file="./data/df_final.csv")
dim(df) # 181 rows and 585 (created an extra ID (=X) variable)

library(plyr)
df <- plyr::rename(df, c("X"="id"))

df_names <- colnames(df)
for(i in df_names) {
  df[[i]] <- as.numeric(df[[i]])
}

df$gender <- as.factor(df$gender)
df$education <- as.factor(df$education)
df$lca_class <- as.factor(df$lca_class)

df$gender <- factor(df$gender, levels = c(1, 2), labels=c("male", "female"))
df$education <- factor(df$education, levels=c(1, 2, 3, 4, 5, 8), 
                       labels=c("primary school", "few secondary sch classes", "secondary school completed", 
                                "vocational training", "university degree", "other"))
df$lca_class <- factor(df$lca_class, levels=c("2", "3", "1"), 
                              labels=c("Disintegrated", "Moderately integrated", "Well-integrated"))


```

<br>

<center> <h1>Demographic information</h1> </center>

### For the total sample:
``` {r, eval=T, echo=F}
table(df$gender) 
table(df$education)
```

The largest part of the sample consists of women. Most of them have secondary education, but about third has a university degree. 
<br>
<br>
Distribution, mean and sd of **age**: 
```{r, eval=T, echo=F}
table(df$age)
#which is: 
mean(df$age, na.rm=T)
sd(df$age, na.rm=T)
```
The distribution of age appears to be approximately normal, with the mean age being 34 years. 
<br>
<br>

### By latent group:
```{r, eval=T, echo=F}
table(df$lca_class) 
t(table(df$lca_class, df$gender))
t(table(df$lca_class, df$education))
library(psych)
t(aggregate(age ~ lca_class, data=df, mean))

```
I will perform the statistical test of independence in the next file to check whether groups differ on demographic variables. 

